
\section{BASILISK explicit energy (\texttt{basilisk})}
\label{sec:basilisk-energy}

This sidechain DBN energy terms allows you to use the BASILISK model
of amino acid sidechain geometry as an \emph{explicit} energy (rather
than the \emph{implicit} term that arises when sampling from the model
and not dividing out the bias). This can of interest when one wishes
to evaluate (or compensate for) the bias introduced by sampling from
the model. It may also be used to evaluate the likelihood (score) of
the side chain geometry. The returned energy is the negative log
likelihood as calculated from the side chain models.

The BASILISK energy term can be used on side chain conformations
in full atomic detail. Using the \texttt{ignore-bb} option, it is
possible to turn off the backbone dependency of the energy.

Both moves use an external model file. Please remember to set the
\texttt{PHAISTOS\_ROOT} environmental variable to avoid problems accessing
these files (see section \ref{sec:locating-data-files})..

For more information on the BASILISK model please consult our 
BASILISK publication \cite{harder2010beyond}.

\optiontitle{Settings}
\begin{optiontable}
  \option{mocapy-dbn-dir}{string} {../data/mocapy\_dbns} {Path to mocapy model file directory.}
  \option{model-filename}{string} {basilisk.dbn} {Filename of the mocapy model.}
  \option{use-cache}	 {bool} {\emph{true}} 	{Use the cached version of the term (should be faster).}
  \option{ignore-bb}	 {bool} {\emph{false}} 	{Whether to omit the backbone conformation (backbone-independent mode)}
\end{optiontable}

\section{COMPAS explicit energy (\texttt{compas})}
\label{sec:compas-energy}

The COMPAS energy term is the pseudo-sidechain atom equivalent to the
full-atom \texttt{basilisk} term. It makes it possible to evaluate an
explicit energy for the COMPAS model (which is normally included
implicitly through the use of the COMPAS move). The COMPAS model is
described in detail in reference \cite{stovgaard2011}.

\optiontitle{Settings}
\begin{optiontable}
  \option{mocapy-dbn-dir}{string} {../data/mocapy\_dbns} {Path to mocapy model file directory.}
  \option{model-filename}{string} {compas.dbn} {Filename of the mocapy model.}
  \option{use-cache}	 {bool} {\emph{true}} 	{Use the cached version of the term (should be faster).}
  \option{ignore-bb}	 {bool} {\emph{false}} 	{Include the backbone conformation for more accurate evaluation.}
\end{optiontable}


